Structural Divergence Analysis

### 2017-02-28 analysis-structure-core.Rmd
cambios 28/02/17

```{r set-file-names}
# set input filenames
## data
reference.fname = file.path(data.dir, paste(family, "_ref.txt", sep = ""))
protein_list.fname = file.path(data.dir, paste(family, "_list.txt", sep = ""))
site.info.fname = file.path(data.dir, paste(p.ref, "_m.da.ca.csv", sep = ""))
min.da.CM.CA.fname = file.path(data.dir, paste(p.ref, "_min.da.CM.ca.csv", sep = ""))

## profiles to compare
dri2.exp.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.exp.norm.dr.squarei.csv", sep = ""))
dri2.strong.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_strong.sel_K.analysis_Keff_out_m.theo.norm.dr.squarei.csv", sep = ""))
dri2.mut.fname = file.path(data.dir,paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.theo.norm.dr.squarei.csv", sep = ""))
```

@@

```{r set-site-info}
# read site information of the reference protein
site.info = read.csv(site.info.fname, header = T, sep = ",")
m.min.da.CM.CA = read.csv(min.da.CM.CA.fname)

@@

>>>>

```{r set-file-names}
# set input filenames
## data
reference.fname = file.path(data.dir, paste(family, "_ref.txt", sep = ""))
protein_list.fname = file.path(data.dir, paste(family, "_list.txt", sep = ""))
site.info.fname = file.path(data.dir, paste(p.ref, "_m.da.ca.csv", sep = ""))
min.da.CM.CA.fname = file.path(data.dir, paste(p.ref, "_min.da.CM.ca.csv", sep = ""))
sum.inv.da.CM.CA.fname = file.path(data.dir, paste(p.ref, "_sum.inv.da.CM.ca.csv", sep = ""))

## profiles to compare
dri2.exp.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.exp.norm.dr.squarei.csv", sep = ""))
dri2.strong.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_strong.sel_K.analysis_Keff_out_m.theo.norm.dr.squarei.csv", sep = ""))
dri2.mut.fname = file.path(data.dir,paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.theo.norm.dr.squarei.csv", sep = ""))

## consurf
score.fname = file.path(data.dir, paste(p.ref, "_consurf.csv", sep = ""))
sim.data.no.sel.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_no.sel_out_df.data.csv", sep = ""))
sim.data.strong.sel.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_strong.sel_out_df.data.csv", sep = ""))
```

@@

```{r set-site-info}
# read site information of the reference protein
site.info = read.csv(site.info.fname, header = T, sep = ",")
m.min.da.CM.CA = read.csv(min.da.CM.CA.fname)
m.sum.inv.da.CM.CA = read.csv(sum.inv.da.CM.CA.fname)

@@

```{r comparison-sum.inv.da}

# comparison of means.z.scores with the sum.inv distance to the functional sites
plot(y = d.exp$mean.z.RSD - d.mut$mean.z.RSD, x = m.sum.inv.da.CM.CA$sum.inv.da.CA[indexes.core], xlab = "sum.inv distance to active site (between CAs)", ylab = "exp.mean.z.RSD - mut.mean.z.RSD")
plot(y = d.exp$mean.z.RSD - d.mut$mean.z.RSD, x = m.sum.inv.da.CM.CA$sum.inv.da.CM[indexes.core], xlab = "sum.inv distance to active site (between CMs)", ylab = "exp.mean.z.RSD - mut.mean.z.RSD")

## calculate correlations between profiles
r.sum.inv.da.CA = cor((d.exp$mean.z.RSD - d.mut$mean.z.RSD), m.sum.inv.da.CM.CA$sum.inv.da.CA[indexes.core])
r.sum.inv.da.CM = cor((d.exp$mean.z.RSD - d.mut$mean.z.RSD), m.sum.inv.da.CM.CA$sum.inv.da.CM[indexes.core])

cor.df = data.frame("comparison" = c("(exp - mut) vs sum.inv.da.CA", "(exp - mut) vs sum.inv.da.CM"), "R" = c(r.sum.inv.da.CA, r.sum.inv.da.CM))

knitr::kable(cor.df, digits = 2, caption = "Correlation between differences in <Z-scores> profiles and the sum.inv distance to the active site")
```

```{r correlation-da-CN} 
## CN core
CN = CN[indexes.core]

## calculate correlations between profiles
r.min.da.CA.CN = cor(CN, m.min.da.CM.CA$min.da.CA[indexes.core])
r.min.da.CM.CN = cor(CN, m.min.da.CM.CA$min.da.CM[indexes.core])

cor.df = data.frame("comparison" = c("CN vs min.da.CA", "CN vs min.da.CM"), "R" = c(r.min.da.CA.CN, r.min.da.CM.CN))

knitr::kable(cor.df, digits = 2, caption = "Correlation between CN and the minimal distance to the active site")

## calculate correlations between profiles
r.sum.inv.da.CA.CN = cor(CN, m.sum.inv.da.CM.CA$sum.inv.da.CA[indexes.core])
r.sum.inv.da.CM.CN = cor(CN, m.sum.inv.da.CM.CA$sum.inv.da.CM[indexes.core])

cor.df = data.frame("comparison" = c("CN vs sum.inv.da.CA", "CN vs sum.inv.da.CM"), "R" = c(r.sum.inv.da.CA.CN, r.sum.inv.da.CM.CN))

knitr::kable(cor.df, digits = 2, caption = "Correlation between CN and the sum.inv distance to the active site")
```

```{r}
# read experimental data obtained from ConsurfDB
consurf = read.csv(score.fname, sep = ";")
score = consurf$SCORE[indexes.core]

# read simulated data
## no sel
sim.data.no.sel = read.csv(sim.data.no.sel.fname)
mut.site.no.sel = sim.data.no.sel$site
p.accept.no.sel = sim.data.no.sel$p.accept

## strong sel
sim.data.strong.sel = read.csv(sim.data.strong.sel.fname)
mut.site.strong.sel = sim.data.strong.sel$site
p.accept.strong.sel = sim.data.strong.sel$p.accept

# calculate de mean.p.accept of simulated data
mean.p.accept.no.sel = c()
mean.p.accept.strong.sel = c()

if (data.dir == "OUT/out_subset_CM_ANM") {
  mut.site.no.sel = mut.site.no.sel - n.sites
  mut.site.strong.sel = mut.site.strong.sel - n.sites
  CN.CM = sim.data.no.sel$CN
  CN = c()

  for (i in (1:n.sites)) {
    CN[i] = CN.CM[which(mut.site.no.sel == i)][1]
  }
  CN = CN[indexes.core]
}

# core n.sites

for (i in (1:n.sites)) {
  p.accept.no.sel.i = p.accept.no.sel[which(mut.site.no.sel == i)]
  mean.p.accept.no.sel.i = mean(p.accept.no.sel.i)
  mean.p.accept.no.sel[i] = mean.p.accept.no.sel.i

  p.accept.strong.sel.i = p.accept.strong.sel[which(mut.site.strong.sel == i)]
  mean.p.accept.strong.sel.i = mean(p.accept.strong.sel.i)
  mean.p.accept.strong.sel[i] = mean.p.accept.strong.sel.i
}
mean.p.accept.no.sel = mean.p.accept.no.sel[indexes.core]
mean.p.accept.strong.sel = mean.p.accept.strong.sel[indexes.core]

# compare exp vs no sel profiles
## plot
par(pty = "s")

plot(x = mean.p.accept.no.sel, y = CN)
plot(x = mean.p.accept.no.sel, y = score)
plot(x = score, y = CN)

## calculate corelations
r.mean.p.accept.no.sel.CN = cor(mean.p.accept.no.sel, CN)
r.mean.p.accept.no.sel.score = cor(mean.p.accept.no.sel, score)
r.score.CN = cor(score, CN)

cor.df = data.frame("comparison" = c("p.accept.no.sel vs CN", "p.accept.no.sel vs score", "score vs CN"),
                    "R" = c(r.mean.p.accept.no.sel.CN , r.mean.p.accept.no.sel.score, r.score.CN))

knitr::kable(cor.df, digits = 2, caption = "Correlation between sequence profiles")

# compare exp vs strong sel profiles
## plot
par(pty = "s")

plot(x = log(mean.p.accept.strong.sel), y = CN)
plot(x = log(mean.p.accept.strong.sel), y = score)
plot(x = score, y = CN)

## calculate corelations
r.mean.p.accept.strong.sel.CN = cor(log(mean.p.accept.strong.sel), CN)
r.mean.p.accept.strong.sel.score = cor(log(mean.p.accept.strong.sel), score)
r.score.CN = cor(score, CN)

cor.df = data.frame("comparison" = c("log(p.accept.strong.sel) vs CN", "log(p.accept.strong.sel) vs score", "score vs CN"),
                    "R" = c(r.mean.p.accept.strong.sel.CN , r.mean.p.accept.strong.sel.score, r.score.CN))

knitr::kable(cor.df, digits = 2, caption = "Correlation between sequence profiles")

```

### 2017-02-28 analysis-structure.Rmd
cambios 28/02/17

### sequence comparison
```{r}
consurf = read.csv(score.fname, sep = ";")
score = consurf$SCORE

sim.data.no.sel = read.csv(sim.data.no.sel.fname)
mut.site.no.sel = sim.data.no.sel$site
p.accept.no.sel = sim.data.no.sel$p.accept

sim.data.strong.sel = read.csv(sim.data.strong.sel.fname)
mut.site.strong.sel = sim.data.strong.sel$site
p.accept.strong.sel = sim.data.strong.sel$p.accept

mean.p.accept.no.sel = c()
mean.p.accept.strong.sel = c()

if (data.dir == "OUT/out_subset_CM_ANM") {
  mut.site.no.sel = mut.site.no.sel - n.sites
  mut.site.strong.sel = mut.site.strong.sel - n.sites
  }

@@

par(pty = "s")

@@

r.mean.p.accept.no.sel.CN = cor(mean.p.accept.no.sel, CN)
r.mean.p.accept.no.sel.score = cor(mean.p.accept.no.sel, score)
r.score.CN = cor(score, CN)

@@

par(pty = "s")

@@

r.mean.p.accept.strong.sel.CN = cor(log(mean.p.accept.strong.sel), CN)
r.mean.p.accept.strong.sel.score = cor(log(mean.p.accept.strong.sel), score)
r.score.CN = cor(score, CN)

>>>>

### sequence comparison
```{r sequence-profiles-comparison}
# read experimental data obtained from ConsurfDB
consurf = read.csv(score.fname, sep = ";")
score = consurf$SCORE

# read simulated data
## no sel
sim.data.no.sel = read.csv(sim.data.no.sel.fname)
mut.site.no.sel = sim.data.no.sel$site
p.accept.no.sel = sim.data.no.sel$p.accept

## strong sel
sim.data.strong.sel = read.csv(sim.data.strong.sel.fname)
mut.site.strong.sel = sim.data.strong.sel$site
p.accept.strong.sel = sim.data.strong.sel$p.accept

# calculate de mean.p.accept of simulated data
mean.p.accept.no.sel = c()
mean.p.accept.strong.sel = c()

if (data.dir == "OUT/out_subset_CM_ANM") {
  mut.site.no.sel = mut.site.no.sel - n.sites
  mut.site.strong.sel = mut.site.strong.sel - n.sites
  CN.CM = sim.data.no.sel$CN
  CN = c()

  for (i in (1:n.sites)) {
    CN[i] = CN.CM[which(mut.site.no.sel == i)][1]
  }
}

@@

# compare exp vs no sel profiles
## plot
par(pty = "s")

@@

## calculate corelations
r.mean.p.accept.no.sel.CN = cor(mean.p.accept.no.sel, CN)
r.mean.p.accept.no.sel.score = cor(mean.p.accept.no.sel, score)
r.score.CN = cor(score, CN)

@@

# compare exp vs strong sel profiles
## plot
par(pty = "s")

@@

## calculate corelations
r.mean.p.accept.strong.sel.CN = cor(log(mean.p.accept.strong.sel), CN)
r.mean.p.accept.strong.sel.score = cor(log(mean.p.accept.strong.sel), score)
r.score.CN = cor(score, CN)
